---
title: "Comparison of Different Name Standardization Approaches"
---

Summary: In this notebook, I look at the association between BNI and longevity. To test robustness of the results, I compared three different BNI indicators:

(1) BNI of original BUNMD names
(2) BNI of standardized name (MPC crosswalk)
(3) BNI of "strengthened" name (MPC crosswalk + manual additions + additional ABE nicknames)


```{r}
library(data.table) ## for working with big data
library(lfe) ## for fixed effects (with lots of FEs)
library(stargazer) ## for regression output tables
library(tidyverse) ## data manipulation and data viz 
library(broom) ## extract regression output
```

## Add standardized names and BNI Incides 

```{r}
## dt fc-server
dt <- fread("~/Downloads/bunmd_v1/bunmd_sib_data.csv")

## dt local 
## dt <- fread("/censoc/data/working_files/bunmd_sib_data.csv")

## read in master nickname file (cleaned)
nicknames_master <- fread("../data/nickname_crosswalk_master.csv") %>%
  mutate(nickname_master = 1) %>% 
  rename(fname_std_master = fname_std)

## join onto dt 
dt <- merge(dt, nicknames_master, all.x = T, by = c("sex", "fname"))
dt[is.na(nickname_master), nickname_master := 0]
dt[is.na(fname_std_master), fname_std_master := fname]

## note: key == "" is in here 35886
dt = dt[byear <= 1920]

# add root bni (master)

bni_root <- read_csv("../data/bni_root_master.csv") %>% 
  rename(bni_root_master = bni_root, n_fname_root_master = n_fname_root)

dt <- left_join(dt, bni_root, by = c("fname_std_master" = "fname_std"))


## add root bni (MPC Crosswalk)

## read in master nickname file (cleaned)
nicknames_mpc <- fread("../data/mpc_nicknames.csv")

## join onto dt 
dt <- merge(dt, nicknames_mpc, all.x = T, by = c("sex", "fname"))
dt[is.na(nickname_mpc), nickname_mpc := 0]
dt[is.na(fname_std), fname_std := fname]

## root bni
bni_root <- read_csv("../data/bni_root.csv") 

## calculate key
dt[sex == 1, nkey_male := .N, by = key]

## add 
dt <- left_join(dt, bni_root, by = c("fname_std"))
```

## run models 

```{r}
dt[, sex_score := mean(sex), by = fname_std_master]

m.bni = felm(death_age ~ bni | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              race == 2 &
              sex == 1 &
            n_fname >= 500,
            data = dt)

m.bni.std = felm(death_age ~ bni_root | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              race == 2 &
              sex == 1 &
            # sex_score < 1  &
            n_fname_root >= 500,
            data = dt)

m.bni.std.strong = felm(death_age ~ bni_root_master | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              race == 2 &
              sex == 1 &
            n_fname_root >= 500,
            data = dt)

stargazer(m.bni, m.bni.std, m.bni.std.strong,
          type = "text", 
           object.names = T)
```

## run models controlling for sex ratio 

```{r}
dt[, sex_score := mean(sex), by = fname]
dt[, sex_score_std := mean(sex), by = fname_std]
dt[, sex_score_std_strong := mean(sex), by = fname_std_master]

m.bni = felm(death_age ~ bni | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              race == 2 &
              sex == 1 &
              sex_score < 1.3  &
              byear %in% c(1900:1920) &
            n_fname >= 500,
            data = dt)

m.bni.std = felm(death_age ~ bni_root | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              race == 2 &
              sex == 1 &
             sex_score_std < 1.3  &
             byear %in% c(1900:1920) &
            n_fname_root >= 500,
            data = dt)

m.bni.std.strong = felm(death_age ~ bni_root_master | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              race == 2 &
              sex_score_std_strong < 1.3 &
              byear %in% c(1900:1920)  &
            n_fname_root_master >= 500,
            data = dt)

stargazer(m.bni, m.bni.std, m.bni.std.strong,
          type = "text", 
           object.names = T)
```

=========================================================================
                                     Dependent variable:                 
                    -----------------------------------------------------
                                          death_age                      
                           (1)               (2)               (3)       
                          m.bni           m.bni.std     m.bni.std.strong 
-------------------------------------------------------------------------
bni                      -0.906*                                         
                         (0.471)                                         
                                                                         
bni_root                                  -0.928**                       
                                           (0.473)                       
                                                                         
bni_root_master                                              -0.641      
                                                             (0.487)     
                                                                         
-------------------------------------------------------------------------
Observations              9,194             9,889             9,977      
R2                        0.697             0.675             0.671      
Adjusted R2               0.234             0.232             0.228      
Residual Std. Error 4.771 (df = 3634) 4.775 (df = 4178) 4.781 (df = 4248)
=========================================================================
Note:                                         *p<0.1; **p<0.05; ***p<0.01


## Northern States Only

```{r}
south_socstate.vec = c(10, 11, 12, 13, 24, 37, 45, 51, 54, 1, 21, 28, 47, 05, 22, 40, 48)*100

dt[, south_socstate := socstate %in% south_socstate.vec]

m.bni = felm(death_age ~ bni | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              south_socstate == FALSE &
              race == 2 &
              sex == 1 &
              sex_score < 1.3  &
              byear %in% c(1900:1920) &
            n_fname >= 500,
            data = dt)

m.bni.std = felm(death_age ~ bni_root | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              south_socstate == FALSE &
              race == 2 &
              sex == 1 &
             sex_score_std < 1.3  &
             byear %in% c(1900:1920) &
            n_fname_root >= 500,
            data = dt)

m.bni.std.strong = felm(death_age ~ bni_root_master | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              south_socstate == FALSE &
              race == 2 &
              sex_score_std_strong < 1.3 &
              byear %in% c(1900:1920)  &
            n_fname_root_master >= 500,
            data = dt)

stargazer(m.bni, m.bni.std, m.bni.std.strong,
          type = "latex", 
           object.names = T)
```

## Minimum Frequency Sensitivity Analysis 

```{r}
min_freqs <- seq(from = 1, to = 3000, by= 100)

results <- list()

for (min in min_freqs) {
  
m.bni = felm(death_age ~ bni | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              race == 2 &
               sex_score < 1.3  &
              sex == 1 &
               byear %in% c(1900:1920) &
            n_fname >= min,
            data = dt)

m.bni.std = felm(death_age ~ bni_root | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              byear %in% c(1900:1920) &
              race == 2 &
              sex == 1 &
              sex_score_std < 1.3  &
            # sex_score < 1  &
            n_fname_root >= min,
            data = dt)

m.bni.std.strong = felm(death_age ~ bni_root_master | key + birth_order + as.factor(byear),
            subset =
              nkey_male %in% 2:5 &
              race == 2 &
              sex == 1 &
              sex_score_std_strong < 1.3  &
              byear %in% c(1900:1920) &
            n_fname_root_master >= min,
            data = dt)

m.bni.df <- tidy(m.bni) %>% 
  mutate(term = "BNI")

m.bni.std.df <- tidy(m.bni.std) %>% 
  mutate(term = "BNI Standardized")

m.bni.std.strong.df <- tidy(m.bni.std.strong) %>% 
  mutate(term = "BNI Standardized (Strong)")

results[[min]] <- bind_rows(m.bni.df, m.bni.std.df, m.bni.std.strong.df) %>% 
  mutate(min_freq = min)

}

results_df <- bind_rows(results)

BNI_gradient <- results_df %>% 
  ggplot() + 
  geom_line(aes(x = min_freq, y = estimate)) + 
  geom_point(aes(x = min_freq, y = estimate)) +
  geom_errorbar(aes(x = min_freq, y = estimate, ymin = estimate - 1.96*std.error, ymax = estimate + 1.96*std.error)) +
  theme_bw(base_size = 20) + 
  geom_hline(yintercept  = 0, linetype = "dashed") + 
  labs(x = "Minimum Frequency",
       y = "BNI Gradient") + 
  facet_wrap(~term)

```




